---
title: "Illustrate error with stan_rdump"
author: "Jacqueline Buros"
date: "12/14/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
if (!requireNamespace("bladderbatch", quietly = TRUE))
  BiocManager::install("bladderbatch")
library(bladderbatch)
library(tidyverse)
library(brms)
```

## Review data

Let's load some example data from the `bladderbatch` package.

```{r load-data}
data(bladderdata, package = 'bladderbatch')
# Get the expression data
edata = exprs(bladderEset)
# Get the pheno data
pdata = pData(bladderEset)
```


Next we convert to a long format for input to brms

```{r}
expr_data <- edata %>%
  as.data.frame() %>%
  tibble::rownames_to_column('gene') %>%
  tidyr::gather(sample_desc, normalized_expression, -gene) %>%
  dplyr::left_join(pdata %>%
                     dplyr::mutate(sample_desc = colnames(edata)),
                   by = 'sample_desc')
```

Note that this dataset comprises expression for 57 samples

```{r}
assertthat::assert_that(
  expr_data %>% dplyr::group_by(sample) %>%
    dplyr::n_groups()
  == 57
  )
```

and 22,283 genes

```{r}
assertthat::assert_that(
  expr_data %>% dplyr::group_by(gene) %>% 
    dplyr::n_groups()
  == 22283
)
```


In this example we will look at only the "Cancer" samples for simplicity

```{r}
training_data <- expr_data %>% 
  dplyr::filter(cancer == 'Cancer') 
```

among these samples, we have 3 outcomes

```{r}
training_data %>%
  dplyr::distinct(outcome)
```

Let's imagine we want to evaluate how gene expression differs in sTCC vs mTCC

```{r}
training_data <- training_data %>%
  dplyr::mutate(is_mtcc = ifelse(stringr::str_detect(outcome, pattern = 'mTCC'), 1, 0))
training_data %>% dplyr::distinct(is_mtcc, outcome)
```

We are left with data for 40 samples. This is too many samples when fitting with brms, so we select a subset for illustration.

```{r}
assertthat::assert_that(
  training_data %>% dplyr::group_by(sample) %>% n_groups() 
  == 40)
training_sample <- training_data %>%
  dplyr::group_by(batch) %>%
  dplyr::distinct(sample_desc) %>%
  dplyr::sample_n(3) %>%
  dplyr::ungroup()
```

## Prepare data for `stan_rdump`

```{r}
# brms inputs (hypothetically)
brms_inputs <- list(formula = brms::bf(normalized_expression ~ 1 + is_mtcc + is_mtcc:gene + (1 | batch) + (1 | gene)),
                    data = training_data %>% dplyr::semi_join(training_sample, by = c('sample_desc', 'batch')),
                    sparse = T,
                    priors = brms::set_prior(horseshoe(df = 3, par_ratio = 0.1), class = 'b'),
                    family = hurdle_gamma)
brms_stan_code <- purrr::invoke(brms::make_stancode, brms_inputs)
brms_stan_data <- purrr::invoke(brms::make_standata, brms_inputs)
```

Write out to file using `stan_rdump`

```{r}
## errors
try(
  rstan::stan_rdump(names(brms_stan_data), file = 'test_rdump.Rds', envir = list2env(brms_stan_data))
  )
```

## Session info 

```{r}
sessionInfo()
```